---
title: "Emil Zonta - Getting and Cleaning Data Course Project"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    theme: zenburn
---

All of my input data is inside a `R/Getting_and_Cleaning_Data_Course_Project/` directory, so I first set my working dir to `R/Getting_and_Cleaning_Data_Course_Project/UCI HAR Dataset` in order to easily load data.

According to the *README.txt* file contained in the downloaded archive, the original dataset is splitted into **training** and **test** data.
Each one again splittein into three files containing respectively:
	1. all of the estimators collected, 
	2. training labels, i.e the activities
	3. subjects who performed the activities

I load the training data. 
```{r}
X_train       <- read.table('train/X_train.txt')
y_train       <- read.table('train/y_train.txt')
subject_train <- read.table('train/subject_train.txt')
```

Same for test data.
```{r}
X_test        <- read.table('test/X_test.txt')
y_test        <- read.table('test/y_test.txt')
subject_test  <- read.table('test/subject_test.txt')
```

Now I can merge the training and the test sets to create one data set.
First of all I merge the three original training data sets with the corresponding three test data sets.
I need to merge them vertically (row binding), so I use the `rbind` function.
```{r}
X_total <- rbind(X_train,X_test)
y_total <- rbind(y_train,y_test)
subject_total <- rbind(subject_train,subject_test)
```

Now I want to merge these columns horizontally, through `cbind`, to obtain a unique data set as requested.
There is no need of using the `merge` function since raw data is correctly sorted.
```{r}
total <- cbind(subject_total,y_total,X_total)
```

Now I read the features labels and transform them into a vector to assign new colnames.
Of course the 1st column of **total** is now the subject, while the 2nd is y hence the activity and the remaining ones are features.
```{r}
features <- read.table('features.txt')
featuresVector <- as.vector(features$V2)
colnames(total) <- c("subject","activity",featuresVector)
```

I just need subject, activity and mean and std values of features.
I use regexps+grep to select them, since I am looking for `*mean()*` and `*std()*` columns.
```{r}
total <- total[, grep('subject|activity|mean\\(\\)|std\\(\\)',colnames(total))]
```

Finally I substitute integer values of activities with descriptive names, as required.
```{r}
activityLabels <- read.table("activity_labels.txt")
total[,2] <- activityLabels[total[,2],2]
```

Last step: create a second, independent tidy data set with the average of each variable for each activity and each subject.
I use the `ddply` function from the `plyr` package.
```{r}
library(plyr)
totalAvg <- ddply(total, .(subject, activity), function(x) colMeans(x[,3:68],na.rm=TRUE))
```

Just printing a portion of this final dataset:
```{r}
library(knitr)
kable(totalAvg[1:18,1:8],digits=3)
```